A Fredholm integral equation approach to computing Stokes eigenvalues in the plane

### 2019-03-11 doc/draft-01-stokes/03analysis.tex
adding section 3.6

\begin{thm}
\label{thm:dlmain}
Suppose that $\Omega$ is a simply connected domain, then the operator
$(\cI - 2\cD_{k} - 2 \cW)$ is not invertible if and only if $k$ is 
Dirichlet eigenvalue for Stokes equation on $\Omega$.
\end{thm}
\begin{proof}
Suppose that $k$ is not a Dirichlet eigenvalue for Stokes equation on
$\Omega$. 
Suppose further that $\bmu$ satisfies
\begin{equation}
(\cI - 2\cD_{k} - 2\cW) \bmu = 0 \, , \label{eq:dlproofrep}
\end{equation}
i.e. $\bmu$ is in the null-space
of $(\cI - 2 \cD_{k} - 2 \cW)$. 
Applying the operator $\cW$ to~\cref{eq:dlproofrep} and 
using~\cref{lem:propnullspacecorr}, we get
\begin{equation}
0 = \cW [(\cI - 2\cD_{k} - 2\cW)\bmu] = -2\cW [\bmu] \, .
\end{equation}
Thus~\cref{eq:dlproofrep} reduces to
\begin{equation}
(\cI - 2 \cD_{k})\bmu = 0\, .
\end{equation}
Suppose now $\bu = -2\bD_{k}[\bmu]$ in $\Omega$.
Then $\bu$ is a solution to the oscillatory Stokes equation in $\Omega$,
and applying~\cref{lem:jump-conds}, we get that the interior
limit of the velocity $\bu^{-} = (\cI - 2\cD_{k}) \bmu = 0$ on $\Gamma$. 
Since $k$ is not a Dirichlet eigenvalue for the Stokes equation on 
$\Omega$, we conclude that $\bu \equiv 0$ in $\Omega$. 
This in particular implies that the interior limit of the surface traction
denoted by $\bt^{-} = 0$ on $\Gamma$. 
Using~\cref{lem:jump-conds} again for the jump in surface traction
for the double layer potential, we conclude that the exterior limit
of the surface traction $\bt^{+} = 0$ on $\Gamma$. 
Note that $\bu$ is a radiating solution of the oscillatory
Sotkes equation in the exterior $E$.
From uniqueness of the Neumann problem in the exterior $E$, we conclude
that $\bu \equiv 0$ in $E$ as well, which in particular implies that
the exterior limit of the velocity $\bu^{+}=0$ on $\Gamma$.   
Using the jump conditions in~\cref{lem:jump-conds} again, we
get that $\bmu = \frac{1}{2}(\bu^{+} - \bu^{-}) = 0$.
Thus $(\cI - 2\cD_{k} - 2\cW)$ is invertible if $k$ is not a Dirichlet
eigenvalue for the Stokes equation on $\Omega$.

@@

From Green's representation for the oscillatory Stokes equation, 
we know that 
  \begin{equation} \label{eq:rep-theorem}
    \bS [\bt](\xx) - \bD[\bu](\xx) = \begin{cases} 
    \bu(\xx) &\quad \xx \in \Omega \, , \\
    0 &\quad \xx \in \Omega^{C} \; .
    \end{cases}
  \end{equation}
Suppose thatw $k$ is Dirichlet eigenvalue for the Stokes equation on $\Omega$
and let $\bu$ denote the corresponding eigenfunction and let $\bt$ denote
the surface traction associated with the eigenfunction.
Note that the Green's representation theorem implies that 
$\cS_{k}[\bt^{-}] = 0$, since $\bu^{-}=0$ on $\Gamma$. 
Applying the Green's theorem to the pair $\bt^{-}, \bu^{-}$ and taking 
its derivative on $\Gamma$ using~\cref{lem:jump-conds}, we get
\begin{equation}
\bt^{-} = (\cDt_{k} + \frac{1}{2} \cI) \bt^{-} \, . 
\end{equation}
Combining these two identities, we get
that
\begin{equation}
(-\cI - 2\cD_{k} -2i\eta \cS_{k}) \bt^{-} = 0 \, .
\end{equation}
As in the proof of~\cref{thm:dlmain}, letting $c = \left< \bt^{-} ,\bnu \right>$, 
it follows that
\begin{equation}
(-\cI - 2\cD_{k} - 2i \eta \cS_{k} - 2\cW) (\bt^{-} - c\bnu) = 0 \, ,
\end{equation}
where $\bt^{-} - c\bnu \neq 0$.
Since $\bt^{-} - c\bnu$ is non-trivial, and $i\eta \cS_{k}, \cW$ are self-adjoint ($i\eta \cS_{k}$ 
is self-adjoint since we consider inner products without conjugation), 
it follows from the Fredholm alternative 
that the operator $\cI - 2\cD(k) -2ik \cS - 2\cW$ is also not invertible.
\end{proof}

\subsection{Computational Fredholm determinants}

>>>>

\begin{thrm}
\label{thm:dlmain}
Suppose that $\Omega$ is a simply connected domain, then the operator
$(\cI - 2\cD_{k} - 2 \cW)$ is not invertible if and only if $k$ is 
Dirichlet eigenvalue for Stokes equation on $\Omega$.
\end{thrm}
\begin{proof}
Suppose that $k$ is not a Dirichlet eigenvalue for Stokes equation on
$\Omega$. 
Suppose further that $\bmu$ satisfies
\begin{equation}
(\cI - 2\cD_{k} - 2\cW) \bmu = 0 \, , \label{eq:dlproofrep}
\end{equation}
i.e. $\bmu$ is in the null-space
of $(\cI - 2 \cD_{k} - 2 \cW)$. 
Applying the operator $\cW$ to~\cref{eq:dlproofrep} and 
using~\cref{lem:propnullspacecorr}, we get
\begin{equation}
0 = \cW [(\cI - 2\cD_{k} - 2\cW)\bmu] = -2\cW [\bmu] \, .
\end{equation}
Thus~\cref{eq:dlproofrep} reduces to
\begin{equation}
(\cI - 2 \cD_{k})\bmu = 0\, .
\end{equation}
Suppose now $\bu = -2\bD_{k}[\bmu]$ in $\Omega$.
Then $\bu$ is a solution to the oscillatory Stokes equation in $\Omega$,
and applying~\cref{lem:jump-conds}, we get that the interior
limit of the velocity $\bu^{-} = (\cI - 2\cD_{k}) \bmu = 0$ on $\Gamma$. 
Since $k$ is not a Dirichlet eigenvalue for the Stokes equation on 
$\Omega$, we conclude that $\bu \equiv 0$ in $\Omega$. 
This in particular implies that the interior limit of the surface traction
denoted by $\bt^{-} = 0$ on $\Gamma$. 
Using~\cref{lem:jump-conds} again for the jump in surface traction
for the double layer potential, we conclude that the exterior limit
of the surface traction $\bt^{+} = 0$ on $\Gamma$. 
Note that $\bu$ is a radiating solution of the oscillatory
Sotkes equation in the exterior $E$.
From uniqueness of the Neumann problem in the exterior $E$, we conclude
that $\bu \equiv 0$ in $E$ as well, which in particular implies that
the exterior limit of the velocity $\bu^{+}=0$ on $\Gamma$.   
Using the jump conditions in~\cref{lem:jump-conds} again, we
get that $\bmu = \frac{1}{2}(\bu^{+} - \bu^{-}) = 0$.
Thus $(\cI - 2\cD_{k} - 2\cW)$ is invertible if $k$ is not a Dirichlet
eigenvalue for the Stokes equation on $\Omega$.

@@

From Green's representation for the oscillatory Stokes equation, 
we know that 
  \begin{equation} \label{eq:rep-theorem}
    \bS [\bt](\xx) - \bD[\bu](\xx) = \begin{cases} 
    \bu(\xx) &\quad \xx \in \Omega \, , \\
    0 &\quad \xx \in \Omega^{C} \; .
    \end{cases}
  \end{equation}
Suppose thatw $k$ is Dirichlet eigenvalue for the Stokes equation on $\Omega$
and let $\bu$ denote the corresponding eigenfunction and let $\bt$ denote
the surface traction associated with the eigenfunction.
Note that the Green's representation theorem implies that 
$\cS_{k}[\bt^{-}] = 0$, since $\bu^{-}=0$ on $\Gamma$. 
Applying the Green's theorem to the pair $\bt^{-}, \bu^{-}$ and taking 
its derivative on $\Gamma$ using~\cref{lem:jump-conds}, we get
\begin{equation}
\bt^{-} = (\cDt_{k} + \frac{1}{2} \cI) \bt^{-} \, . 
\end{equation}
Combining these two identities, we get
that
\begin{equation}
(-\cI - 2\cD_{k} -2i\eta \cS_{k}) \bt^{-} = 0 \, .
\end{equation}
As in the proof of~\cref{thm:dlmain}, letting $c = \left< \bt^{-} ,\bnu \right>$, 
it follows that
\begin{equation}
(-\cI - 2\cD_{k} - 2i \eta \cS_{k} - 2\cW) (\bt^{-} - c\bnu) = 0 \, ,
\end{equation}
where $\bt^{-} - c\bnu \neq 0$.
Since $\bt^{-} - c\bnu$ is non-trivial, and $i\eta \cS_{k}, \cW$ are self-adjoint ($i\eta \cS_{k}$ 
is self-adjoint since we consider inner products without conjugation), 
it follows from the Fredholm alternative 
that the operator $\cI - 2\cD(k) -2i\eta \cS - 2\cW$ is also not invertible.
\end{proof}

\subsection{Computational Fredholm determinants}
Having established the equivalence between the Dirichlet eigenvalues
for Stokes equation and the invertibility of
the operator $\cI - 2\cD_{k} - 2\cW$ on simply connected domains, we show how 
the Fredholm determinant can be as a computational tool for detecting
the non-invertibility of $\cI - 2\cD_{k} - 2\cW$.

First, we briefly review Nystr\"{o}m discretizations for integral operators
with weakly singular kernels.
Suppose that $\gamma(t):[0,L] \to \Gamma$ is an arc-length
paramterization of $\Gamma$, where $L$ is the length of the
curve $\Gamma$.
Using this parametrization, the integral equation
$(\cI - 2\cD_{k} - 2\cW)\bmu = \ff$, 
can be reformulated as a system of integral equations on 
$\bL^{2}[0,L]$ of the form 
\begin{equation}
\bmu(x) + \int_{0}^{L} \bK(x,y) \bmu(y) dy = \ff(x) \, , 
\end{equation}
where $\bK(x,y)$ is $O(|x-y|^2 \log{|x-y|})$ as $x\to y$, and
is analytic away from $x\neq y$.

Let $K$ denote the integral operator with kernel $\bK$, i.e.
\begin{equation}
K[\mu](\bx) = \int_{0}^{L} \bK(x,y) \bmu(y) dy \, .
\end{equation}

Suppose that $[0,L]$ is discretized using $N$ discretization
nodes $x_{j}$, $j=1,2,\ldots N$, along with an
the quadrature rule, $Q_{N}(x,x_{j,N})$, $j=1,2,\ldots N$.

Let $K_{N}$ be the finite dimensional Nystr\"{o}m approximation of
the integral operator whose kernel is $\bK(x,y)$ defined for all
continuous functions $\phi$ via the formula:
\begin{equation}
K^{N}[\phi](\bx) = \sum_{j=1}^{N} Q^{N}(x,x^{N}_{j})  
\bK(x,x_{j}^{N}) \phi(x,x_{j}^{N}) \, .
\end{equation}

The linear system corresponding to the Nystr\"{o}m discretization
of $\bK$ denoted by $\tilde{K}^{N}$ is then given by
\begin{equation}
\tilde{K}^{N}_{i,j} = Q^{N}(x_{i}^{N}, x_{j}^{N}) \bK(x_{i}^{N}, x_{j}^{N})
\quad i,j=1,2\ldots N \, .
\end{equation}

\begin{definition}
The quadrature rule $Q_{N}$ is considered
to be $p$th order accurate if there exists a $C$ independent of $N$
such that for all $\bmu \in C^{p}$, 
\begin{equation}
\lim_{N\to \infty} |(K_{N} - K)\bmu|_{\bL^{2}([0,L])} \leq C_{\bmu} N^{-p} \, .
\end{equation}
\end{definition}

\begin{remark}
There are many sufficient conditions on
the choice of $x_{j,N}, $Q_{N}(x,x_{j,N})$ $j=1,2\ldots,N$, $N \in mathbb{N}$
for which the corresponding Nystr\"{o}m 
\end{remark}

given byi
be the finite rank operator 
The Nystr\"{o}m discretized linear system corresponding to the operator
$\bK$ is the matrix whose entries are given by
\begin{equation}
K_{N} = Q_{N}(x_{i,N},x_{j,N}) \bK(x_{i,N},x_{j,N}) \, .
\end{equation}

The standard Nystr\"{o}m procedure is to solve the integral equation
\begin{equation}
\bmu_{i,N} + \sum_{j=1}^{N} Q_N(x_{i,N},x_{j,N} \bK(x_{i,N},x_{j,N})
\bmu_{j,N} = \ff(\bx_{i,N}) \, ,
\end{equation}
for the unknowns $\bmu_{i,N}$, along with the natural interpolant
\begin{equation}
\bmu(\bx) = \ff(\bx) - \sum_{j=1}^{N} Q_N(x,x_{j,N}) \mu_{j,N} \, .
\end{equation}

\begin{equation}
\end{equation}
%Suppose that the quadrature rule $Q_{N}$ satisfies
%the following two additional constraints:
%a) 
%\begin{equation}
%\sum_{j=1}^{N} |Q_{N}(x,x_{j,N})| \leq C \, ,
%\end{equation}
%where $C$ is a constant independent of $x$ and $N$, and b) that
%\begin{equation}
%\lim_{x\to y} \sup_{n \in \mathbb{N}} \sum_{j=1}^{N} 
%|Q_{N}(x,x_{j,N}) - Q_{N}(y,x_{j,N})| \to 0 \, .
%\end{equation}

Suppose that the domain $[0,L]$, is partitioned into $N$ 
panels of equal length, where each panel is discretized 
using order $p$ discretization nodes (e.g.

Let $\cL = -2\cD_{k} - 2\cW$, and let $\cJ_{1}(X)$ denote the 
space of trace class operators on the Hilbert Space $X$.
Suppose the boundary $\Gamma$ is parametrized into $N$ panels
of equal length.

\begin{definition}
\end{definition}

The following theorem describes how 
the determinant of a Nystr\"{o}m
discretized second kind Fredholm equation where the integral
kernel is weakly-singular converges to $0$ if and only if
the operator is not invertible.

\begin{thrm}
\label{thm:mainconvfreddet}
Suppose that $A:X \to X$ is a compact operator, and that $X$ is a
Hilbert space
\end{thrm}

\begin{proof}
See~\cite{zhao2015robust}, theorem 4.3 for a detailed discussion. 
\end{proof}

\begin{remark}
The same analysis does not carry through for the operator
$\cI - 2 \cD_{k} - 2i\eta \cS_{k} - 2\cW$, 
since it is not a trace class operator.
For brevity, let $\cM_{k} = -2\cD_{k} - 2i\eta \cS_{k} - 2\cW$.
The operator $\cM$ is in $\cJ_{2}(\bL^{2}(\Gamma))$ where
$\cJ_{2}(\bL^{2}(\Gamma))$ is the space of Hilbert-Schmidt operators
on $\bL^{2}(\Gamma)$ (the singular values of the operator are square
summable, as opposed to being summable).
Thus the Fredholm determinant of $\cI + \cM$
is not necessarily convergent. 
However, as noted in~\cite{zhao2015robust}, the convergence i
result~\cref{thm:mainconvfreddet}
should be true up to a logarithmic factor in the rate
of convergence, since the singular values of the operator
$\cM_{k}$ decay like $\frac{1}{n}$, and the
Fredholm determinant divergers logarithmically. 
Let $M_{N,k}$ denote the Nystr\"{o}m discretized operator 
corresponding to $M$, where $N$ is the number of discretization points
on the boundary $\Gamma$.
One way of proving such a result would be to construct a definition 
of a determinant for Hilbert-Schmidt operators based on the 
Diximer trace of operators as opposed to the Fredholm determinant
and prove analogous results for this case.
Instead, in~\cref{subsec:convannulus}, we demonstrate that
that $\text{det} (I - M^{N,k}) \to 0$ as $N \to \infty$ if and only 
if $k$ is a Dirichlet eigenvalue for the Stokes operator, and that
the convergence to $0$ is exponential at the eigenvalues even in 
this case.
\end{remark}
